---
title: "PHW251 Problem Set 3"
author: "Answer Key"
format: pdf
editor: visual
---

Render this file from the `problem_set_3` folder so the relative paths to `../data` and `table_replicate.png` resolve.

\newpage

### Question 1

Create a data frame with three columns: ID (1, 2, 3), NAME ("Pam", "Jim", "Dwight"), and AGE (40, NA, 48).

```{r office_df}
office_df <- data.frame(
  ID = c(1, 2, 3),
  NAME = c("Pam", "Jim", "Dwight"),
  AGE = c(40, NA, 48)
)

office_df
```

\newpage

### Question 2

Find the length, typeof, and class of the data frame from Question 1.

```{r office_values}
length(office_df)
typeof(office_df)
class(office_df)
```

A data frame is a list of columns, so `length()` is 3 (the number of columns), `typeof()` is `"list"`, and `class()` is `"data.frame"`.

\newpage

### Question 3

Create a data frame and a tibble that match `table_replicate.png`: columns `data_id`, `gender`, and `temperature`.

```{r tibble}
library(tibble)

covid_df <- data.frame(
  data_id = c(101, 102, 103, 104, 105),
  gender = c("female", "male", "non-binary", "male", NA),
  temperature = c(98, 97.3, 101.1, 97.5, 99.6)
)

covid_tib <- tibble(
  data_id = c(101, 102, 103, 104, 105),
  gender = c("female", "male", "non-binary", "male", NA),
  temperature = c(98, 97.3, 101.1, 97.5, 99.6)
)

covid_df
covid_tib
```

\newpage

### Question 4

What are the key differences between data frames and tibbles? Why are tibbles preferable?

Data frames and tibbles both store rectangular data, but tibbles change a few defaults that make them easier to work with:

-   Tibbles do not use row names.
-   Tibbles keep column names as you write them, including names with spaces or names that start with a number. Data frames will alter some of those names.
-   A length-1 vector in a tibble is recycled to every row. You do not have to repeat a constant yourself.
-   Inside `tibble()`, you can refer to a column you are creating in the same call.
-   Printing a tibble shows the column types and only the first few rows, instead of dumping the whole table.
-   Subsetting a tibble with `[, ]` still returns a tibble. The same subset of a data frame can drop down to a vector.

Tibbles are preferable because that behavior is more predictable, the printout is easier to read, and the rest of the tidyverse is built to work with them.

\newpage

### Question 5

Add a `results` column to the tibble from Question 3.

-   101 = NEGATIVE
-   102 = POSITIVE
-   103 = NEGATIVE
-   104 = NEGATIVE
-   105 = NEGATIVE

```{r covid_results}
covid_tib$results <- c("NEGATIVE", "POSITIVE", "NEGATIVE",
                       "NEGATIVE", "NEGATIVE")
covid_tib
```

\newpage

### Question 6

Subject 102's temperature is 98.3, not 97.3.

```{r fix_covid_results}
covid_tib$temperature[covid_tib$data_id == 102] <- 98.3
covid_tib
```

\newpage

### Question 7

Load `stds-by-disease-county-year-sex.csv` from the data folder. The first three rows are not part of the table, so skip them.

```{r import_stds}
stds <- read.csv(file = "../data/stds-by-disease-county-year-sex.csv",
                  skip = 3)
head(stds)
```

\newpage

### Question 8

How many rows?

```{r}
nrow(stds)
```

How many columns?

```{r}
ncol(stds)
```

What are the column names?

```{r}
colnames(stds)
```

What are the column types?

```{r}
sapply(stds, class)
```

There are 9558 rows and 6 columns: Disease, County, Year, Sex, Cases, and Population. Year, Cases, and Population are integers. The other three columns are character.

\newpage

### Question 9

Keep the rows for 2015-2018 with `which()`. The subset should be 2124 rows by 6 columns.

```{r which}
stds_recent <- stds[which(stds$Year >= 2015 & stds$Year <= 2018), ]
dim(stds_recent)
```

\newpage

### Question 10

Export the 2015-2018 subset as a csv in the data folder, then read it back.

```{r export}
write.csv(stds_recent,
          file = "../data/stds_2015_2018.csv",
          row.names = FALSE)

stds_check <- read.csv(file = "../data/stds_2015_2018.csv")
head(stds_check)
dim(stds_check)
```

\newpage

### Question 11

`unique()` on County should return 59 counties. Then subset to one county.

```{r unique}
counties <- unique(stds$County)
counties
length(counties)
```

```{r county_subset}
alameda <- stds[stds$County == "Alameda", ]
head(alameda)
```

\newpage

### Question 12

In the county subset, add `rate = (Cases / Population) * 100,000`.

```{r rate}
alameda$rate <- (alameda$Cases / alameda$Population) * 100000
head(alameda)
```

\newpage

### Question 13

Reply to the classmate whose import code is not a repeatable example. Include at least three specific suggestions.

Three changes that would make this a repeatable example:

1.  Delete `rm(list = ls())`. Clearing the whole environment is not part of the example, and it can wipe objects the other person needs in order to see the problem.
2.  Load the package before using `read_csv()`. A repeatable example has to include `library(readr)` or `library(tidyverse)`, because another computer will not already have that function attached.
3.  Do not depend on a local file the reader does not have. Paste a few rows with `dput()`, or use a small built-in data set, and include the exact error message. The arguments `header` and `row.names` also belong to `read.csv()`, not `read_csv()`, so the call should use the arguments for the function that is actually being run.